What We Eval in the Shadows

### 2021-04-15 paper/main.tex
Iterate on Usage of eval

\mypara{Variable lookup and values.}
This corresponds to minimized expressions \c{X}, \c{V}. We also include here lookup to slots in a list or data frame, \ie minimized expression \c{\$}.
It is the predominant usage of \eval in Base, through function \c{match.args}. In CRAN, it correlates with more exotic environments passed to \eval. \todo{THis is with a V in practice.}
The dominant site in numbers of calls in CRAN, in the custom object system of \emph{ggplot2}, \c{ggproto}, also evaluates a variable \c{`\_inherit`} in the definition environment of the class. It leads to a large diversity of environments, many being package environments or package environments wrapped in \c{new}.
\todo{Explain why they use eval and not get... because of \c{::}}

@@

\todo{Show an actual example with \$}
Slot lookup happens very often to evaluate some of the elements of a call list returned by \c{match.call}.
For instance in the \emph{pwrFDR} package:
\begin{lstlisting}
m <- match.call()
...
average.power <- m$average.power <- eval(m$average.power, sys.parent())
\end{lstlisting}

Here, \c{m\$average.power}, depending on the shape of the arguments to the surrounding function,  can be resolved to a variable, a more complex expression, or a number, leading to several possible minimized expressions, including \c{X} and \c{V}.

@@

\mypara{Capturing a call.}  A common use case for \eval is to be combined with \c{match.call}.
\c{match.call} walks up the call stack, captures the code that invoked the
currently executing function, and returns it as an unevaluated expression that includes the call function and its unevaluated arguments. The
pattern is to transform a call to some function \c{f} into a call to \c{g} with
some arguments retained and others modified.  \todo{or change the target
function. Both are described as "patterns" in the R documentation.} As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the rewritten call
is evaluated in the same environment the original call was.  \begin{lstlisting}
tvcglm <- function(formula, data, family, ...) { k <- match.call() k[[1]] <-
as.name("tvcm") if ("fit" %in% names(list(...))) warning("'fit' ignored.")
k$fit <- "glm" eval.parent(k) } \end{lstlisting} This pattern is recognizable
by the fact that the expression is a call and the target environment is that of
the parent.

@@

\todo{Put a simpler example.}

@@

@@

@@

\eval is a powerful tool and can be sometimes with functions with more specialized goals. For instance, variable lookup in any environment can be performed with \c{get}, and assignment in any environment, with \c{assign}. Building a call and executing can be done with \c{do.call}. Nevertheless, \eval shines when expressions passed to it can be of many different types. % And it is shorter to type, and no need to remember all the specialized functions.

>>>>

\mypara{Variable lookup and values.}
This corresponds to minimized expressions \c{X}, \c{V}. We also include here lookup to slots in a list or data frame, \ie minimized expression \c{\$}. Variable lookup leads to a large diversity of environments, many being package environments or package environments wrapped in \c{new}.

It is the predominant usage of \eval in Base, through function \c{match.args} (with \eval on minimized expression \c{V}). In CRAN, it correlates with more exotic environments passed to \eval. 
The dominant site in numbers of calls in CRAN, is located in \c{ggproto} in the custom object system of \emph{ggplot2}. It evaluates a variable \c{`\_inherit`}, which represents the parent of the class, in the definition environment of the class. \eval is used here to account for parents scoped in a namespace with \c{::}, which is actually a function call in R.

@@

\todo{Show an actual example with \$}

@@

\mypara{Capturing a call.}  A common use case for \eval is to be combined with \c{match.call}.
\c{match.call} walks up the call stack, captures the code that invoked the
currently executing function, and returns it as an unevaluated expression that includes the call function and its unevaluated arguments. The
pattern is to transform a call to some function \c{f} into a call to \c{g} with
some arguments retained and others modified.  \todo{or change the target
function. Both are described as "patterns" in the R documentation.} As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the rewritten call
is evaluated in the same environment the original call was.  \begin{lstlisting}
tvcglm <- function(formula, data, family, ...) { k <- match.call() k[[1]] <-
as.name("tvcm") if ("fit" %in% names(list(...))) warning("'fit' ignored.")
k$fit <- "glm" eval.parent(k) } \end{lstlisting} This pattern is recognizable
by the fact that the expression is a call and the target environment is that of
the parent. \todo{Remove that example}

@@

\todo{Show also an example with \c{substitute}}

@@

  \todo{Put a simpler example, again with building a plot}

@@

Using \c{parse}, \ie converting strings to R expressions, opens to a more flexible code generation, as the building blocks which are going to be concatenated together, do not need to be valid R expressions. On the contrary, combining expression created with \c{substitute} or \c{quote} requires more discipline. In Kaggle, which features less experienced programmers,  code generation for plots is widespread, and uses \c{parse}.

@@

\eval is a powerful tool and can be sometimes replaced with functions with more specialized goals. For instance, variable lookup in any environment can be performed with \c{get}, and assignment in any environment, with \c{assign}. Building a call and executing can be done with \c{do.call}. Nevertheless, \eval shines when expressions passed to it can be of many different types. % And it is shorter to type, and no need to remember all the specialized functions.
